%!TEX root = Thesis.tex

\chapter*{Acknowledgements}
\label{ch:acknowledgements}

As always there are numerous people that have helped me over the years.  My supervisor, Prof. John Spencer, thank you for your guidance, advice and support, even coming in to University after a major earthquake to rescue our reaction. Dr. Matthias Lein, thank you for being an excellent supervisor while John was on research and study leave and for all of your assistance with DFT.  Thanks to both of your for reading through various sections while on holiday.

The past and present members of the JLS research group: Almas, Brad, Chris, David, Kathryn, Lia, Rosie, Sarah and Teresa.  Thank you for getting excited with me over the minutiae of NMR spectra, and for all of the good times and commiserations that we've shared.  The other postgraduate students in SCPS, particularly my office mates in AM204, thanks for your friendships over the years.

The general staff of SCPS are amazingly supportive towards the research in the school and many things would be impossible without them.  Particular thanks for the lab technicians Teresa, Jamie-Ann and Jackie, for letting me borrow glassware and other equipment.  Thanks to Peter Northcote, Ian Vorster and John Ryan for helping with NMR.  I am grateful to Rob Keyzers and Yinrong Lu for help with mass spectrometry and to Jan Wikaira, Christ Fitchett and Matt Polson for X-ray crystallography.  

Thank you to the members of Disarray, and all of my non-chemistry friends for giving me something else to think about when I just needed a break.  My family have been extremely supportive and caring, as they always are.  Finally to my husband, Stephen, these last 10 years have been wonderful, thank you for all of your love and understanding especially over these last few months.

I would also like to acknowledge Victoria University of Wellington for funding through the PhD, submission, and Curtis-Gordon scholarships.  
%
%\begin{itemize}
%\item{John and Matthias}
%\item{Lab techs, Teresa, Jamie, Shekira, Jackie}
%\item{NMR, Peter, Ian, John Ryan}
%\item{Mass spec, Rob, Ian, Yinrong}
%\item{Crystallography, Jan, Chris, Matt?}
%\item{Lab group, Rosie, Sarah, Teresa, Kathryn, Brad, Chris, David, Almas}
%\item{Stephen}
%\end{itemize}